% !TeX spellcheck = en_US
% !TeX root = ../Tom_Sandmann-master_thesis
\chapter{SAKURA-G} \label{chp: SAKURA-G}
\lettrine[lhang = 0.4, findent=-30pt, lines=4]{\textbf{
		\initfamily \fontsize{20mm}{20mm} \selectfont T
		\normalfont}}{he}
SAKURA-G board consists of two integrated Spartan-6 FPGAs.
One of these Spartan FPGAs serves as the main security circuit (XC6SLX75-2CSG484C), while the other one serves as the controller (XC6SLX9-2CSG225C). 
The main Spartan contains the actual cryptographic hardware design.
The control Spartan is used to control the main board by changing specific signals received by the main board (e.g. signals indicating whether encryption/decryption has to be performed, or whether the internal state has to be reset). 
To deploy a hardware design on our FPGA, we need to somehow transfer this design to the board.
This requires the hardware design to be processed in a specific way:
% https://www.xilinx.com/itp/xilinx10/isehelp/ise_c_implement_fpga_design.htm
\begin{enumerate}
	\item \textbf{Synthesis}: the abstract description of our hardware design (written in for example VHDL or Verilog) is turned into a design implementation of logic gates and lookup tables (LUTs), digital signal processors (DSPs), BRAMs and other elements.
	
	\item \textbf{Mapping, Place and Route}: the structures identified in the previous step are mapped to FPGA elements.
	These components are then routed and the appropriate signals are connected.
	
	\item \textbf{Program file}: a file is generated that can be transferred to the FPGA. Depending on the file format, this file either gets flashed to the flash memory on the SAKURA-G board, or is stored in the FPGA non-persistent memory.
\end{enumerate}
%

\section{Constraints}
Constraints are used to guide the design tool on how specific parts of the design should be treated. There are two types of constraints:
%
\begin{itemize}
	\item \textbf{Synthesis constraints} are used by the synthesis tool to optimize specific parts of the hardware description language (HDL) code. 
	They can be either embedded directly within the VHDL/Verilog code or specified in an external synthesis constraint file.
	
	\item \textbf{Implementation constraints} are instructions passed to FPGA implementation tools that specify mapping, placement, timing and other guidelines followed by the implementation tool while processing a FPGA design. These constraints are generally placed in a User Constraint File (\texttt{UCF}). Examples of these constraints are \texttt{LOC} (placement) and \texttt{PERIOD} (timing) constraints. In the hardware design we consider, the majority of the constraints (if not all) are \texttt{LOC} constraints. \texttt{LOC} constraints define where a design element can be placed within a FPGA. 
\end{itemize}
%

\section{JTAG}
JTAG was a standard initially developed by IEEE to solve issues with electronically manufactured boards. 
It is a standard used to verify designs and circuit board after they have been manufactured.
In our case, it is used as a programming, debug and probing port. 
The JTAG programmer/debugger is attached to both the JTAG port and the micro-USB port, which makes it ready to use. 
We use the JTAG programmer to load hardware designs of the control and main FPGAs to the Spartan FPGAs.
In general, there are two file formats that can be loaded to a FPGA:
%
\begin{itemize}
	\item \texttt{BIT} file: A \texttt{.BIT} file is a raw storage of the programming bits for the FPGA.
	It can be loaded to the FPGA via JTAG using for example \texttt{iMPACT} (a Xilinx specific utility).
	This file format is primarily used for testing a hardware design. The reason for this is, when the board loses its power, the design is ``lost''.
	
	\item \texttt{MCS} file: A \texttt{.MSC} file is flashed to flash memory, which means that the contents are not lost when power is lost. The \texttt{.MSC} file can be flashed to flash-memory via JTAG using \texttt{iMPACT}. On power-up, specific configuration signals are used to load the program to the board.
\end{itemize}
%

\input{sections/interface_with_board}
\input{sections/ip_blocks}
\input{sections/reading_writing_internal_registers}
\input{sections/capturing_power_traces}








