% !TeX spellcheck = en_US
% !TeX root = ../Tom_Sandmann-master_thesis
\section{\texorpdfstring{The curve: \fourq}{The Curve FourQ}}
A new complete twisted Edwards curve was introduced in \cite{costello2015fourq}.
They work over the following quadratic extension field:
%
\begin{align*}
\mathbb{F}_{p^2} := \mathbb{F}_p(i), \text{ where } p := 2^{127} - 1 \text{ and } i^2 = -1
\end{align*}
%
%To explain what an extension field is, we first explain what a subfield is.
%If we have a field $L$, a subfield of this field $L$ is a subset $K$ that is also a field with respect to the inherited field operations from $L$.
%In a similar way, a subfield is a subset that contains the element 1.
%In addition, subfield is closed under the addition, subtraction and multiplication operations, and also in taking the inverse of a non-zero element of $L$.
They define $\mathcal{E}$ to be the following twisted Edwards curve:
\begin{align*}
\mathcal{E} \slash \mathbb{F}_{p^2} : -x^2 + y^2 = 1 + dx^2 y^2
\end{align*}
where $p = 2^{127} - 1$ and $d$ is the following non-square in $\mathbb{F}_{p^2}$:
%
\begin{align*}
d := 125317048443780598345676279555970305165 \cdot i + 4205857648805777768770
\end{align*}
%
The set of rational points $\mathbb{F}_{p^2}$ together with $\mathcal{E}$ form a group, in which the neutral element is $\mathcal{O} = (0, 1)$ and the inverse of a point $(x, y)$ is $(-x, y)$.
This curve, which is named \fourq, is based on the use of $\mathbb{Q}$-curve reductions as described in \cite{smith2016mathbb}. 
The group $\mathbb{F}_{p^2}$-rational points on $\mathcal{E}$ is denoted by $\mathcal{E}(\mathbb{F}_{p^2})$.
Every point in this group is an element of the product set $\mathbb{F}_{p^2}$, which means that it is a sequence (or in this case a tuple) of 2 elements of $\mathbb{F}_{p}$ (as denoted by the power of $p$).
Every element in the field $\mathbb{F}_{p^2}$ can therefore be seen as complex number.
Group operations follow the elementary operations as defined for complex numbers. 
During the remainder of this chapter, we work in the cryptographic group $\mathcal{E}(\mathbb{F}_{p^2}) \left[ N \right]$, where $N$ is a 246-bit prime (which is fixed according to \cite[Equation 2]{costello2015fourq}).

\section{\texorpdfstring{{\fourqs} scalar multiplication}{FourQ's scalar multiplication}} \label{sec: FourQ's scalar multiplication}
{\fourqs} scalar multiplication routine can be seen in \Cref{algo: FourQ's scalar multiplication}.
First, the endomorphisms for point $P$ are computed.
Next, the lookup table $T$ is precomputed.
This uses the endomorphisms computed earlier.
After decomposing and recoding the scalar, the algorithm continues to execute the main loop.
This is where the point doublings and additions of the algorithm take place.
Each of previously described steps are explained in more detail in the upcoming subsections.
%
\begin{algorithm}
	\algorithmicrequire Point $P \in \mathcal{E}(\mathbb{F}_{p^2}) \left[ N \right]$ and integer scalar $m \in [0, 2^{256})$.\\
	\algorithmicensure $[m]P$.
	%	
	\begin{algorithmic}[1]
		\Statex \textbf{Compute endomorphisms}
		\State Compute $\phi(P), \psi(P)$ and $\psi(\phi(P))$ as described in \Cref{subsec: The endomorphisms psi and phi}.
		\Statex \textbf{Precompute lookup table}
		\State Compute $T[u] = P + [u_0] \phi(P) + [u_1] \psi(P) + [u_2] \psi(\phi(P))$ for $u = (u_2, u_1, u_0)_2$ in $0 \le u \le 7$. \label{lst:fourq_scalar_mult:precompute table}
		\Statex \textbf{Scalar Decomposition}
		\State Decompose $m$ into the multiscalar $(a_1, a_2, a_3, a_4)$ as described in \Cref{subsec: Scalar decomposition}.
		\Statex \textbf{Scalar recoding}
		\State Recode $(a_1, a_2, a_3, a_4)$ into $(d_{64}, \ldots , d_0)$ and $(m_{64}, \ldots, m_0)$ as described in \Cref{subsec: Scalar multiplication}. Write $s_i = 1$ if $m_i = -1$ and $s_i = -1$ if $m_i = 0$.
		\Statex \textbf{Main loop:}
		\State $Q = s_{64} \cdot T[d_{64}]$ \label{lst:fourq scalar mult:initial assignment}
		\For{$i = 63$ \textbf{downto} $0$}
			\State $Q = [2]Q$				\label{lst:fourq scalar mult:double oper}
			\State $Q = Q + s_i \cdot T[d_i]$ \label{lst:fourq scalar mult:add oper}
		\EndFor
		\State \textbf{Return} $Q$
	\end{algorithmic}
	%
	\captionof{algorithm}{{\fourqs} scalar multiplication on $\mathcal{E}(\mathbb{F}_{p^2}) \left[ N \right]$ \cite{costello2015fourq}.}
	\label{algo: FourQ's scalar multiplication}
\end{algorithm}
%

\input{subsections/fourq_endomorphisms}
\input{subsections/fourq_decomposition}
\input{subsections/fourq_multiplication}
\input{subsections/fourq_results}